\emph{Mesh networks} are a special kind of \emph{Ad-Hoc networks}:
usually they are more structured and tend to present semi-permanent
nodes. However the structure of the network itself is not fixed. Nodes
may join and leave and links between nodes may change during
time. This peculiar characteristic requires a distributed routing
strategy capable of handling such a scenario.

There are two main family of protocols developed to handle routing in
\emph{mesh networks}: \emph{pro-active} and \emph{reactive}.
The former maintain fresh lists of destinations and their routes by
periodically distributing routing tables throughout the network, the
latter finds a route on demand by flooding the network with \emph{route
request packets}.

Thinking about the \emph{pro-active} family a couple of questions
comes to our minds:
\begin{itemize}
\item   How the overhead of the routing protocols affect the network
        performances?
\item   How long is the convergence time on the event of a sudden link
        removal?
\end{itemize}

In this report we shall analyze these aspects. We'll show a series of
results obtained through a some experiments on two \emph{pro-active} mesh
routing protocols: \batman\cite{bib:BATMAN} and \olsr\cite{bib:OLSR}.

\subsection{About \batman}

    Each host in a \batman\ mesh network produces on a
    regular basis an \emph{Originator Message}, which is sent in
    broadcast and gets received by all directly reachable neighbors. Every
    \emph{Originator Message} contains, among other things, a
    \emph{Sequence Number} and a \emph{TTL}; it is used to
    inform the network of the presence of the node.

    Each node keeps, for any destination host, a sliding window of
    recent sequence numbers coming from the various available routes. The
    window gets moved as new Sequence Numbers show up, thus discarding
    obsolete ones. The preferred
    route for a specific destination is given by the host which has
    forwarded, in the current window,  the most number of \emph{Originator
    Messages} generated by the destination.

    Any node is in charge of re-broadcasting an \emph{Originator Message}
    if:
    \begin{itemize}
    \item   The Sequence Number hasn't been seen yet;
    \item   The message's TTL hasn't been expired.
    \end{itemize}

    The suggested period for the flooding operation is of 1 second, thus
    in an environment configured with default values this is the actual
    frequency of the flooding operation.

\subsection{About \olsr}

    The protocol is far more complex than \batman\ and it's best suited for
    extended and complex networks. Such complexity boils down to
    each host periodically sending a portion of the local \emph{link-state}
    table. The details are beyond the purposes of this report, since
    they're not affecting that much the actual network interaction. What
    really concerns us is the software timing:
    \begin{itemize}
    \item   A probing for the direct neighbors is achieved by means of a
            periodically broadcasted \emph{Hello} message. The
            \Const{HELLO\_INTERVAL} constant, determining the broadcasting
            period, is of 2 seconds;
    \item   Some of the direct neighbors are used as relay nodes, and are
            targeted with messages of type \emph{Topology Control}.
            The \Const{TC\_INTERVAL} constant schedules a direct
            communication every 5 seconds.
    \end{itemize}

\subsection{Experiments with meshed networks}
    As briefly stated above we've executed two kind of tests: the
    first one aims to measure the worsening of the network
    performance in presence of routing protocols with respect to a static
    routing configuration; in the second test we measure how
    responsive is each protocol in case of changes in the network
    topology.

    The needed data has been retrieved by using some tools:
    \begin{itemize}
    \item   Data about latency has been obtained by means of the
            \emph{ping} tool;
    \item   Throughput has been measured with
            \netperf\cite{bib:NetPerf};
    \item   Statistics about protocol responsiveness for \batman\ and
            \olsr\ have been extrapolated by parsing the output of the
            two softwares respectively;
    \item   Complementary data has been extrapolated
            from \emph{Whireshark} logs.
    \end{itemize}

\subsection{The testbed}
   The testbed has been composed by four laptops equipped with
    \emph{Ubuntu 10.10 GNU/Linux} operating systems. Table~\ref{tab:Tools}
    shows a list of the programs we used.

    \begin{table}
        \centering
        \begin{tabular}{lll}
            \toprule
            Software  & Version & Note \\
            \midrule
            \netperf  & 2.4.4-5ubuntu2 \\
            \batman   & 0.3.2-5 & userspace version \\
            \olsr     & 0.5.6-r7-1 \\
            Iptables  & 1.4.4-2ubuntu3 \\
            Wireshark & 1.2.11-4build0.10.10.10 \\
            \bottomrule
        \end{tabular}
        \caption{List of softwares used for this experiments}
        \label{tab:Tools}
    \end{table}

    Depending on the tests, the network topologies have been arranged
    specifically, so far each section of this report will be provided with
    a paragraph describing it.
